Simple connectome inference from partial correlation statistics in calcium imaging

### 2014-06-27 article.tex
break 80 char in appendix

Besides filters
$f_1$ and $f_2$ respectively defined by equations \ref{eq:symetric-median} and \ref{eq:weighted-asymetric-median}, we also consider two other low-pass
filters:

@@

The magnification can also be optimized by adjusting differently the effect depending on the global activity and the low-pass filter used. In that sense, we introduce the following function \begin{align}
 w^*(x^{t}_i) &= {(x^{t}_i + 1 )^{\left (1 + \frac{1}{\sum_{j} x^{t}_j}\right )}}^{m}
\end{align}
where $m$ is a piecewise linear function depending on the low-pass filter and on the range of the global activity. We refer to our implementation\footnote{available on \url{https://github.com/asutera/kaggle-connectomics}.} for the values of the parameters and especially the parameter $m$.

@@

The heuristic is based on the following observation. The rise of fluorescence
of a neuron indicates its activation. If another neuron is activated after a
slight delay, this could have been caused by the activation of the first
neuron and therefore highlight a (direct) directed relationship between the
two neurons. With that in mind, the heuristic principle is to count the number
of such events occurred between neurons $i$ and $j$ for consecutive time
steps. It can be done by counting the number of times that a transition from
neuron $i$ to neuron $j$ for two consecutive time steps, or mathematically we
compute for each pair $(i,j)$

\begin{align}
s_{ij} = \sum_{t=1}^{T - 1} \mathbb{1}((x_j^{t+1} - x_i^t) \in \left[\phi_1, \phi_2\right])
\end{align}
where $\phi_1$ and $\phi_2$ are parameters of the heuristic.

Two issues may be spotted. First, this heuristic is not able to distinguish a
directed relationships if both neurons are activated in the same time step and
may wrongly characterize indirect relationships because of the time
resolution. Second, it may be necessary to take into account $s_{ji}$ to
estimate the quality of the pairwise coefficient $s_{i,j}$. Indeed, if both
coefficients are similar ($s_{ij} \sim s_{ji}$) the directivity can not be
inferred since the order of fluorescent rising is not well defined. On the
contrary, different coefficients $s_{ij}$ and $s_{ji}$ mean that an order of
fluorescent rising is well established and observable. There we choose to
compute a differential pairwise coefficient $z_{ij}$ defined by the following
equation

\begin{align}
z_{ij} = \sum_{t=1}^{T - 1}
    \mathbb{1}((x_j^{t+1} - x_i^t) \in \left[\phi_1, \phi_2 \right]) -
    \mathbb{1}((x_i^{t+1} - x_j^t) \in \left[\phi_1, \phi_2 \right]),
\end{align}
where $T$ is the number of time steps, $\phi_1$ and $\phi_2$ are parameters.

>>>>

Besides filters $f_1$ and $f_2$ respectively defined by equations \ref{eq
:symetric-median} and \ref{eq:weighted-asymetric-median}, we also consider two
other low-pass filters:

@@

The magnification can also be optimized by adjusting differently the effect
depending on the global activity and the low-pass filter used. In that sense,
we introduce the following function
\begin{align}
 w^*(x^{t}_i) &= {(x^{t}_i + 1 )^{\left (1 + \frac{1}{\sum_{j} x^{t}_j}\right )}}^{m}
\end{align}
where $m$ is a piecewise linear function depending on the low-pass filter and
on the range of the global activity. We refer to our
implementation\footnote{available on \url{https://github.com/asutera/kaggle-
connectomics}.} for the values of the parameters and especially the parameter
$m$.

@@

The heuristic is based on the following observation. The rise of fluorescence
of a neuron indicates its activation. If another neuron is activated after a
slight delay, this could have been caused by the activation of the first
neuron and therefore highlight a (direct) directed relationship between the
two neurons. With that in mind, the heuristic principle is to count the number
of such events occurred between neurons $i$ and $j$ for consecutive time
steps. It can be done by counting the number of times that a transition from
neuron $i$ to neuron $j$ for two consecutive time steps, or mathematically we
compute for each pair $(i,j)$
\begin{align}
s_{ij} = \sum_{t=1}^{T - 1} \mathbb{1}((x_j^{t+1} - x_i^t) \in \left[\phi_1, \phi_2\right])
\end{align}
where $\phi_1$ and $\phi_2$ are parameters of the heuristic.

Two issues may be spotted. First, this heuristic is not able to distinguish a
directed relationships if both neurons are activated in the same time step and
may wrongly characterize indirect relationships because of the time
resolution. Second, it may be necessary to take into account $s_{ji}$ to
estimate the quality of the pairwise coefficient $s_{i,j}$. Indeed, if both
coefficients are similar ($s_{ij} \sim s_{ji}$) the directivity can not be
inferred since the order of fluorescent rising is not well defined. On the
contrary, different coefficients $s_{ij}$ and $s_{ji}$ mean that an order of
fluorescent rising is well established and observable. There we choose to
compute a differential pairwise coefficient $z_{ij}$ defined by the following
equation
\begin{align}
z_{ij} = \sum_{t=1}^{T - 1}
    \mathbb{1}((x_j^{t+1} - x_i^t) \in \left[\phi_1, \phi_2 \right]) -
    \mathbb{1}((x_i^{t+1} - x_j^t) \in \left[\phi_1, \phi_2 \right]),
\end{align}
where $T$ is the number of time steps, $\phi_1$ and $\phi_2$ are parameters.
